\Large \bf Lissy: Experimenting with on-chain order books

### 2021-09-13 LaTeX-LNCS-2022/sections/paper.tex
small tweaks
fixed the mindmap branch colors
hooks for improvements

\textbf{Contributions.} In this paper, we experiment with on-chain markets to understand in detail if they remain infeasible on Ethereum and what the limiting factors are. Some highlights from our research include answering the following questions:
\begin{itemize}
\item What type of exchange has the best properties on balance? (A call market.)
\item How many orders can be processed on-chain? (Upper-bounded by 152 per block.)
\item How much efficiency can be squeezed from diligently choosing the best data structures? (Somewhat limited: turn 38 trades into 152.)
\item To what extent can we mitigate front-running attacks? (Almost entirely.)
\item Can we stop the exchange's storage footprint on Ethereum from bloating? (Yes, but it is so expensive that it is not worth it.)
\item Are on-chain order books feasible on layer 2? (Yes! Optimistic roll-ups reduce gas costs by 99.88\%.)
\item Which aspects of Ethereum were encountered that required deeper than surface-level knowledge to navigate? (Optimizing gas refunds, Solidity is not truly object-oriented, miner extractable value (MEV) can be leveraged for good, \textblue{add more?})
\item How hard is an on-chain exchange to regulate? (The design leaves almost no regulatory hooks beyond miners (and sequencers on layer 2).)
\end{itemize}

@@

For our experiments and measurements, we implement an exchange from scratch. \cm is a call market that will open for a specified period of time during which it will accept a capped number of orders (\eg 100 orders---parameterized so that all orders can be processed), and these orders are added to a priority queue (discussed in Section~\ref{sec:pq} ). Our vision is the market would be open for a very short period of time, close, and then reopen immediately (\eg every other block). \cm is open source and written in 336 lines (SLOC) of Solidity plus the priority queue (\eg we implement 5 variants, each around 300 SLOC). We tested it with the Mocha testing framework using Truffle~\cite{TruffleO71:online} on Ganache-CLI~\cite{GanacheT25:online} to obtain our performance metrics. Once deployed, the bytecode of \cm is 10,812 bytes plus the constructor code (6,400 bytes) which is not stored. The Solidity source code for \cm and Truffle test files are available in a GitHub repository.\footnote{Github: Link removed for anonymity.} We have also deployed \cm on Ethereum's testnet Rinkeby with flattened (single file) source code of just the \cm base class and priority queue implementations. It is visible and can be interacted with here: \href{https://rinkeby.etherscan.io/address/0x0d91de29c531d074853a5cef7cf9dfeb9c6ec4e0}{[etherscan.io]}. We cross-checked for vulnerabilities with \textit{Slither}\footnote{https://github.com/crytic/slither} and \textit{SmartCheck}\footnote{https://tool.smartdec.net} and it only fails some `informational' warnings that are intentional design choices (\eg a costly loop). All measurements assume a block gas limit of $11\,741\,495$ and 1 gas $=$ 56 Gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Table~\ref{tab:cm_functions} summarizes \cm's primary operations.

@@

\begin{figure}[t]
\begin{floatrow}
\ffigbox{%
\includegraphics[width=6cm]{fig/2.png}%
}{%
  \caption{\scriptsize{Gas costs for enqueuing 50 random integers into five priority queue variants. For the x-axis, a value of 9 indicates it is the 9th integer entered in the priority queue. The y-axis is the cost of enqueuing in gas}. \label{fig:random_insertion}}%

@@

In our \cm variant on Arbitrum, the validators do all computations (both enqueuing and dequeuing) so we choose to use a heap with dynamic array for our priority queue, which balances the expense of both operations. Heaps are 32\% more efficient than linked lists for submitting orders and 29\% less efficient for closing. Recall that without a roll-up, such a priority queue can only match 38 pairs at a cost of 5,372,679 gas. Table~\ref{tab:arbitrum_performance} shows that 38 pairs cost only 6,569 in L1 gas (a 99.88\% savings). This is the cost of submitting the \texttt{closeMarket()} transaction to the inbox to be recorded, which is 103 bytes of calldata. Most importantly, \texttt{closeMarket()} always costs 6,569 even as the number of trades increase from 38 pairs to thousands or millions of pairs. Of course, as the number of trades increase, the work for the validators on L2 increases, as measured in ArbGas. The price of ArbGas in Gwei is not well-established but is anticipated to be relatively cheap.  Arbitrum also reduces the costs for traders to submit an order: from 207,932 to 6,917 in L1 gas.

>>>>

\textbf{Contributions.} In this paper, we experiment with on-chain markets to understand in detail if they remain infeasible on Ethereum and what the limiting factors are. Some highlights from our research include answering the following questions:
\begin{itemize}
\item What type of exchange has the best properties on balance? (A call market.)
\item How many orders can be processed on-chain? (Upper-bounded by 152 per block.)
\item How much efficiency can be squeezed from diligently choosing the best data structures? (Somewhat limited: turn 38 trades into 152.)
\item To what extent can we mitigate front-running attacks? (Almost entirely.)
\item Can we stop the exchange's storage footprint on Ethereum from bloating? (Yes, but it is so expensive that it is not worth it.)
\item Are on-chain order books feasible on layer 2? (Yes! Optimistic roll-ups reduce gas costs by 99.88\%.)
\item Which aspects of Ethereum were encountered that required deeper than surface-level knowledge to navigate? (Optimizing gas refunds, Solidity is not truly object-oriented, miner extractable value (MEV) can be leveraged for good, and bridging assets for layer 2.)
\item How hard is an on-chain exchange to regulate? (The design leaves almost no regulatory hooks beyond miners (and sequencers on layer 2).)
\end{itemize}

@@

For our experiments and measurements, we implement an exchange from scratch. \cm is a call market that will open for a specified period of time during which it will accept a capped number of orders (\eg 100 orders---parameterized so that all orders can be processed), and these orders are added to a priority queue (discussed in Section~\ref{sec:pq} ). Our vision is the market would be open for a very short period of time, close, and then reopen immediately (\eg every other block). \cm is open source and written in 336 lines (SLOC) of Solidity plus the priority queue (\eg we implement 5 variants, each around 300 SLOC). We tested it with the Mocha testing framework using Truffle~\cite{TruffleO71:online} on Ganache-CLI~\cite{GanacheT25:online} to obtain our performance metrics. Once deployed, the bytecode of \cm is 10,812 bytes plus the constructor code (6,400 bytes) which is not stored. The Solidity source code for \cm and Truffle test files are available in a GitHub repository.\footnote{Github: Link removed for anonymity.} We have also deployed \cm on Ethereum's testnet Rinkeby with flattened (single file) source code of just the \cm base class and priority queue implementations. It is visible and can be interacted with here: \href{https://rinkeby.etherscan.io/address/0x0d91de29c531d074853a5cef7cf9dfeb9c6ec4e0}{[etherscan.io]}. We cross-checked for vulnerabilities with \textit{Slither}\footnote{https://github.com/crytic/slither} and \textit{SmartCheck}\footnote{https://tool.smartdec.net} and it only fails some `informational' warnings that are intentional design choices (\eg a costly loop). All measurements assume a block gas limit of $11\,741\,495$ and 1 gas $=$ 56 Gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Table~\ref{tab:cm_functions} summarizes \cm's primary operations.\

\textblue{Add call market basics.}

@@

\begin{figure}[t]
\begin{floatrow}
\ffigbox{%
\includegraphics[width=6cm]{fig/2.png}%
}{%
  \caption{\scriptsize{\textblue{RESIZE THIS} Gas costs for enqueuing 50 random integers into five priority queue variants. For the x-axis, a value of 9 indicates it is the 9th integer entered in the priority queue. The y-axis is the cost of enqueuing in gas}. \label{fig:random_insertion}}%

@@

In our \cm variant on Arbitrum, the validators do all computations (both enqueuing and dequeuing) so we choose to use a heap with dynamic array for our priority queue, which balances the expense of both operations. Heaps are 32\% more efficient than linked lists for submitting orders and 29\% less efficient for closing. Recall that without a roll-up, such a priority queue can only match 38 pairs at a cost of 5,372,679 gas. Table~\ref{tab:arbitrum_performance} shows that 38 pairs cost only 6,569 in L1 gas (a 99.88\% savings). This is the cost of submitting the \texttt{closeMarket()} transaction to the inbox to be recorded, which is 103 bytes of calldata. Most importantly, recording \texttt{closeMarket()} in the inbox will alway cost around 6,569 even as the number of trades increase from 38 pairs to thousands or millions of pairs. Of course, as the number of trades increase, the work for the validators on L2 increases, as measured in ArbGas. The price of ArbGas in Gwei is not well-established but is anticipated to be relatively cheap.  Arbitrum also reduces the costs for traders to submit an order: from 207,932 to 6,917 in L1 gas. In Appendix~\ref{app:rollup}, the full interaction is shown in Figure~\ref{fig:lissyl2}, which illustrate how traders interact with \cm on Arbitrum including bridges, inboxes, sequencers and validators.

### 2021-09-13 LaTeX-LNCS-2022/tables/eval2.tex
hooks for improvements

% = = = = = = = = = = = =%
On-Chain Dealers
& Quote-driven decentralized exchange trades from inventory with public pricing rule (\eg Uniswap v3)
&  Non-custodial \newline
Instant settlement \newline
Moderate performance \newline
Fair execution 
&  Unconventional \newline
Impermanent loss \newline
High price slippage \newline
Intrinsic fees \newline
Front-running attacks \newline
Limited trading strategies 
\\ 
\hline

% = = = = = = = = = = = =%
On-chain Order-Driven Exchanges 
& Order-driven decentralized exchange executes trades between buyers and sellers (\eg \cm)
&  Conventional \newline
Non-custodial \newline
Instant settlement \newline
Low price slippage \newline
Fair execution \newline
Verbose trading strategies \newline
Front-running is mitigable 
& Very low performance \newline
\\ 
\hline

>>>>

% = = = = = = = = = = = =%
On-Chain Dealers
& Quote-driven decentralized exchange trades from inventory with public pricing rule (\eg Uniswap v3)
&  Non-custodial \newline
Instant trading \newline
Moderate performance \newline
Fair execution 
&  Unconventional \newline
Impermanent loss \newline
High price slippage \newline
Intrinsic fees \newline
Front-running attacks \newline
Limited trading strategies \newline
Hard to regulate
\\ 
\hline

% = = = = = = = = = = = =%
On-chain Order-Driven Exchanges 
& Order-driven decentralized exchange executes trades between buyers and sellers (\eg \cm)
&  Conventional \newline
Non-custodial \newline
Low price slippage \newline
Fair execution \newline
Verbose trading strategies \newline
Front-running is mitigable 
& Very low performance \newline
Hard to regulate
\\ 
\hline
